Supporting Information: Thermal sensitivity across forest vertical profiles: patterns, mechanisms, and ecological implications

### 2021-09-30 vertical-thermal-review.Rmd
edit ecology section

While it is clear that drought sensitivity increases with crown height though forest vertical gradients, and hence with DBH, it remains unclear how growth sensitivity to high $T_{air}$, independent of drought, varies across this gradient.
Because VPD increases rapidly with $T_{air}$, high $T_{air}$ is often associated with atmospheric drought, and trees will exhibit drought responses even when even when soil moisture remains high (**REFS**).
However, under conditions that are humid enough for canopy trees can maintain high $g_s$, we do not necessarily expect their photosynthesis (see section 4.2) or woody growth to exhibit higher sensitivity to $T_{air}$ than their understory counterparts. 
In eight mesic forests across the northeast United States, tree-ring records showed steep growth declines of understory trees at higher $T_{air}$, contrasting with the responses of canopy trees [Fig `r fig_treerings`, @rollinsonClimateSensitivityUnderstory2020].
Additional research is required to understand the mechanisms underlying these intriguing differences, and to disentangle size-related tree growth responses to hot-wet versus hot-dry conditions.

>>>>

While it is clear that drought sensitivity increases with crown height though forest vertical gradients, and hence with DBH, it remains unclear how growth sensitivity to high $T_{air}$, independent of drought, varies across this gradient.
Because VPD increases rapidly with $T_{air}$, high $T_{air}$ is often associated with atmospheric drought, likely explaining negative growth responses -- particularly among larger trees -- even if soil moisture remains high [@novickIncreasingImportanceAtmospheric], or when precipitation is statistically accounted for Fig. `r fig_treerings`,@anderson-teixeiraJointEffectsClimate2021].
However, under conditions that are humid enough that canopy trees can maintain high $g_s$, we do not necessarily expect their photosynthesis (see section 4.2) or woody growth to exhibit higher sensitivity to $T_{air}$ than their understory counterparts. 
In eight mesic forests across the northeast United States, tree-ring records showed steep growth declines of understory trees at higher $T_{air}$, contrasting with the responses of canopy trees [Fig `r fig_treerings`, @rollinsonClimateSensitivityUnderstory2020].
Additional research is required to understand the mechanisms underlying these intriguing differences, and to disentangle size-related tree growth responses to hot-wet versus hot-dry conditions.
